A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-30 paper/main.tex
edits after full read through

\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time
}
\shorttitle{Merger Timescales in TNG100}

@@

\begin{abstract}
    % Statement
    % Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles. % could remove this line 
    % Problem 
    The merger rates of high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively, both observationally and theoretically. 
    Since low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) are comparatively under studied, it is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems.
    However, it is unclear if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift. 
    % Our solution
    Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently, 
    %and converted from one system to another 
    %by scaling 
    % GB added below
    wherein separation criteria are scaled
    by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass major pairs as a function of cosmic time, and compare to the merger timescales of similarly-selected high-mass major pairs. 
    Using our framework to select pairs via a scaled separation criterion leads to equivalent merger timescales for low- and high-mass systems at redshifts $z=0-6$. 
    % What are the timescales? 
    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$), leads to timescales that differ by up to $\sim1\Gyr$ between low- and high-mass pairs. 
    % Why it matters 
    Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system. 
    As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
Galaxy merger rates provide an important pathway for tests of hierarchical assembly from$\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}. 
In the era of future surveys with JWST, Rubin, and Roman, these tests will be extended to higher redshifts and lower mass scales than were previously accessible~\citep{Gardner2006,Spergel2015,Robertson2019a,Robertson2019b,Behroozi2020,Martin2022}.

At higher redshift ($z>\kc{xx}$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
However, there is currently no framework for comparing merger rates for low-mass and high-mass galaxies as a function of time, and tensions between predictions from theory and observational studies exist in the literature at low masses~\citep[see e.g. Fig.~10 in][]{RG2015}.

@@

% 4 - Comparing the merger rates of high mass and low mass pairs is important why?
%     - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
%     - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes

% 6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
%     - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact

7 - In addition, our previous work showed that even the evolution of the fractions of high and low mass pairs can't be robustly studied and compared if even the selection criteria for the pairs are off. 
    - This means that the selection of pairs for pair fractions, merger fractions, and for merger timescales, are ALL going to be impacted by the selection criteria that's used. 
    - This is also a problem in observations already, since different samples will have different completeness down to different projected separations, and with different mass ratios, etc. 
    - O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate

8 - Fortunately, our previous work showed that using scaled criteria for pair selection as a function of mass and redshift allows you to fairly compare samples. 
    - This motivated us to study the merger fraction and merger timescales of our galaxy pair sample from the first study, to see how merger timescales might ALSO depend on the selection criteria that are used to pick the pair sample.

9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.  we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.

%%% OUTLINE
In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the TNG100 simulation. 
We examine the evolution of the number of pairs and merger fraction of our sample as a function of redshift in Sec.~\ref{sec:pairprops}.
In Sec.~\ref{sec:results}, we present our findings on the separation and redshift dependence of the merger timescales of low-mass and high-mass pairs, and study the impact of physical and scaled separation criteria for pair selection. 
Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across different redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.

\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated, but physically associated, as in \citet{Chamberlain2024}.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.

@@

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass 
% GB added M_G
($\rm M_{G}$)\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm M_{G}=1\times 10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Requiring pairs to belong to the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.

We require that subhalos constituting a pair meet a minimum subhalo mass ($\Mhalo$) criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For TNG100, this ensures that subhalos are resolved into over 
$\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
For each subhalo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.

@@

Primary subhalos are defined as the subhalo with the highest assigned stellar mass ($M_{*1}$) in their FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$. 
Our sample of major pairs then consists of all pairs of primary and secondary subhalos with 
\begin{align*} 
\mbox{\textbf{low mass primaries:}}&\, 10^{8}< \rm M_{*1} < 5\times10^{9} \Msun \\ 
\mbox{\textbf{high mass primaries:}}&\, 5\times 10^{9}< \rm M_{*1} < 10^{11} \Msun\\
\mbox{\textbf{stellar mass ratio:}}&\,      
    M_{*2}/M_{*1} > 1/4.
\end{align*}
A primary or secondary subhalo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.

A subhalo can belong to different pairs at different redshifts. 
For example, the primary of a pair that merges at $z=2$ can be selected with a different secondary at $z=1$, constituting a new pair. 
More detail regarding the uniqueness of pairs and orbits is discussed in~\ref{sec:methods-unique}.

@@

\subsection{Mergers} \label{subsec:mergers}
Here we describe how subhalo mergers are identified. 
Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers, especially in cases where subhalos and galaxies have different centers of mass~\citep[see e.g.,]{RG2015}. % might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \subfind{} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
The \sublink{} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.

We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \sublink{} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one.

@@

In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs. If the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.} 
The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo. 
Even amongst galaxies of the same approximate stellar mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger (triangles).

For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than that of their primary. 
Because all of our mergers occur when the subhalos are in the same FoF group by definition, in these cases, we use the full orbit starting when the subhalos are first identified as members of the same FoF group through when they merge, as described in Section \ref{subsec:mergers}.
This definition is more robust for considering the full interaction timescales of merging pairs.
%the post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same. }
% a variety of different orbits

@@

The bottom panel shows the orbits of ``Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.

\subsubsection{Uniqueness of orbits}
\label{sec:methods-unique}
Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists. However, we only want to keep a single instance of any given orbit in our catalog so we can track the merger timescale of the pair.

@@

\section{Pair Sample Properties}\label{sec:pairprops}
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
    We only count the number of orbits ``at" a given redshift if the pairs have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
    \caption{The fraction of low-mass (green) and high-mass (pink) post-infall orbits that result in a merger before z=0 as a function of redshift. 
    The horizontal line shows a merger fraction of 0.9, or 90\%. 
    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
    The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Pairs}\label{sec:pairprops-num}
We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.

Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass pairs as a function of redshift from $z=0-6$. 
Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $z=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to zero at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).

As more subhalos experience first infall into a FoF, the number of pairs increases with decreasing redshift. 
On the other hand, as mergers occur, the number of pairs decreases with decreasing redshift, as pairs that have merged are no longer members of the pairs sample.
% The difference between high and low mass
The number of low-mass pairs continues to increase from $z=6$ until around $z=2$, at which point the number of added pairs at a given redshift slows. 
The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.

The number of high-mass pairs continues increasing until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the new pairs being added at each redshift.

@@

\subsection{Merger Fraction}\label{sec:pairprops-frac}
We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
The merger fraction for both mass ranges declines to zero at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.

>>>>

\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time}
\shorttitle{Merger Timescales in TNG100}

@@

\begin{abstract}
    % Statement
    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, nor if they evolve similarly with redshift. 
    % Our solution
    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time, and explore how different selection criteria impact the mass and redshift dependence of merger timescales. 
    % What are the timescales? 
    We find that the merger timescales for low- and high-mass pairs are the same from $z=0-6$ in the absence of any separation criteria.
    We then present a physically-motivated framework for selecting pairs via a scaled separation criteria, wherein pair separations are scaled by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 \Rvir$).  
    Applying these scaled separation criteria recovers equivalent merger timescales for both mass scales at all redshifts.
    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$. 
    % Importance
    As a result, applying the same merger timescales to physical separation-selected pairs will lead to a bias that systematically over-predicts low-mass galaxy merger rates. 
    % could replace the above line with something softer suggesting the application of scaled selection criteria for studies of merger properties of pairs at different mass scales and redshifts, or something like the below:    
    % Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system. 
    % if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.

\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}
Galaxy merger rates provide an important pathway for tests of hierarchical assembly from $\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}. 
In the era of future surveys with JWST, Rubin, and Roman, these tests will be extended to higher redshifts and lower mass scales than were previously accessible~\citep{Gardner2006,Spergel2015,Robertson2019a,Robertson2019b,Behroozi2020,Martin2022}.

At higher redshift ($z\gtrsim1$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2012,Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
However, there is currently no framework for comparing merger rates for low-mass and high-mass galaxies as a function of time, and tensions between predictions from theory and observational studies exist in the literature at low masses~\citep[see e.g. Fig.~10 in][]{RG2015}.

@@

In our previous work, we show that the redshift evolution of pair fractions of isolated low-mass and high-mass pairs in Illustris TNG100 differ significantly, particularly at $z<3$ where pair fractions of low-mass pairs decrease up to $75\%$ between $z=3$ and $z=0$, while high-mass pair fractions peak at $z=0$~\citep{Chamberlain2024}. 
We suggested that differences in the merger timescales of low-mass and high-mass pairs might be the cause of the pair fraction evolution differences.
In our following analysis, we will investigate this question directly and answer whether merger timescales are responsible for these pair fraction differences.

Additionally, we found in \citet{Chamberlain2024} that recovering the pair fraction differences seen between the two mass scales is sensitive to the selection criteria adopted during pair selection.
Physical separation cuts applied equivalently to low-mass and high-mass pairs at all redshifts eliminate the ability to distinguish the behavior of the true pair fraction evolution (see their Figure~5), even for separation cuts as large as $\rsep<300\kpc$. 
Alternatively, when separation criteria vary with the mass and redshift of each system, in particular when the separation is scaled by the virial radius $\Rvir$ of the pair's FoF group halo, the ability to distinguish the underlying pair fraction behavior was recovered for scaled separations as low as $\rsep < 0.5\,\Rvir$ (see their Figure~4). 
Employing scaled separation criteria then permits the equivalent comparison of pair fractions between low-mass and high-mass pairs, and robust comparisons across redshifts from $z=0-4$. 
Since merger rate estimates are derived from close pair fractions and merger timescales, the results from our previous work imply that careful consideration of separation criteria is required for merger timescale and merger rate studies at different mass scales and redshifts as well.

In order to investigate the impact of pair selection criteria on merger timescales specifically, we will track the pairs selected in \chambe{} forwards and backwards in time to construct orbits for each pair.
We will then study the merger timescales of isolated low-mass and high-mass pairs for the same set of physical and scaled separation criteria used in \chambe{} to determine whether merger timescales of isolated pairs are robust to the selection criteria used to determine the pair samples.

%%% OUTLINE
In this paper, we aim to extend the framework for pair selection criteria presented in \chambe{}.
%to examine the merger timescales that result from these various separation criteria.
In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the Illustris TNG100 simulation. 
We examine the redshift evolution of the number of pairs and merger fraction of our sample in Sec.~\ref{sec:pairprops}.
In Sec.~\ref{sec:results}, we present our findings on the separation and redshift dependence of the merger timescales of low-mass and high-mass pairs, and study the impact of physical and scaled separation criteria for pair selection. 
Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.

\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015} from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major low-mass and high-mass pairs (stellar mass ratio $1/4 \leq M_{*2}/M_{*1}\leq 1$) that are isolated, but physically associated, as in \citet{Chamberlain2024}.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.

@@

At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass $\MG$
% GB added M_G
\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass. This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated. The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.} 
\begin{align*}
        \mbox{\textbf{low mass:}}&\,\rm \MG = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
        \mbox{\textbf{high mass:}}&\, \rm \MG =1\times 10^{12}- 6.5\times10^{12}\,\Msun.
\end{align*}
Requiring pairs to belong to the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.

We require the subhalos that constitute a pair to meet a minimum subhalo mass, \Mhalo, criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For TNG100, this ensures that subhalos are resolved into over 
$\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
For each subhalo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.

@@

Primary subhalos are defined as the subhalo with the highest assigned stellar mass, $M_{*1}$ in the FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$. 
Our sample of major pairs then consists of all pairs of primary and secondary subhalos with 
\begin{align*} 
\mbox{\textbf{low mass primaries:}}&\, 10^{8}< M_{*1} < 5\times10^{9} \Msun \\ 
\mbox{\textbf{high mass primaries:}}&\, 5\times 10^{9}< M_{*1} < 10^{11} \Msun\\
\mbox{\textbf{stellar mass ratio:}}&\,      
    M_{*2}/M_{*1} > 1/4.
\end{align*}
A primary or secondary subhalo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
A subhalo can belong to different pairs at different redshifts. 
For example, the primary of a pair that merges at $z=2$ can be selected with a different secondary at $z=1$, constituting a new pair. 
More detail regarding the uniqueness of pairs and orbits is discussed in~\ref{sec:methods-unique}.

@@

\subsection{Mergers} \label{subsec:mergers}
Here we describe how mergers are identified. 
Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers~\citep[see e.g.,][]{RG2015,Patton2024}, especially in cases where dark matter subhalos and galaxies have different centers of mass, as is the case in our own Galaxy~\citep{Gomez2015,GaravitoCamargo2019,Petersen2021,Chamberlain2023}. % might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \subfind{} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
The \sublink{} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.

We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when each merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \sublink{} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one.

@@

In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift. We use the definition of scaled separation $\rsep/\Rvir$ from~\cite{Chamberlain2024}, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs. If the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.} 
The virial radius of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled separation" is, by construction, a function of mass and redshift.
%, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected from \chambe{} at $z=1.5$. 
The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit before the secondary and primary ever share a FoF group halo. 
Even amongst galaxies of the same approximate stellar mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger (triangles).

For some pairs, the secondary subhalos can experience first infall, then after one pericentric passage can return to large distances where they are temporarily assigned a different FoF group halo than that of their primary. 
In these cases, we use the full orbit beginning at first infall, including the sections of the orbit where the secondary is assigned to another FoF group temporarily,  through to merger as described in Section \ref{subsec:mergers}.
This definition is more robust for considering the full interaction timescales of merging pairs.
%the post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same. }
% a variety of different orbits

@@

The bottom panel shows the orbits of ``Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of selected orbits.

\subsubsection{Uniqueness of orbits}
\label{sec:methods-unique}
Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists. However, we only want to keep a single instance of any given orbit in our catalog, since the orbits with the longest timescales will be selected many times which would skew our data to longer merger timescales.

@@

\section{Pair Sample Properties}\label{sec:pairprops}
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of unique low-mass (green) and high-mass (pink) pairs in our orbit catalog as a function of redshift. Orbits are only defined between first infall and the merger redshift (for merging pairs) or $z=0$ (for non-merging pairs). 
    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers), while dashed lines show the number of pairs at a given redshift that will merge before $z=0$.
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{"total"}
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
    \caption{The fraction of low-mass (green) and high-mass (pink) post-infall orbits that result in a merger before z=0 as a function of redshift. 
    The horizontal line shows a merger fraction of 0.9, or 90\%. 
    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
    The sharp decline of the merger fraction to zero at low redshift is a non-physical feature of the simulation ending at $z=0$.
    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Pairs}\label{sec:pairprops-num}
We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.

Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass pairs as a function of redshift from $z=0-6$. 
Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while high-mass pairs (pink solid line) are most numerous between $z=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to zero at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).

As subhalos experience first infall into a FoF group, the number of pairs increases. 
On the other hand, mergers cause a decrease in the number of pairs.
% The difference between high and low mass
The number of low-mass pairs increases from $\sim600$ pairs at $z=6$ to $\sim6,600$ at $z=2$. 
The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.

The number of high-mass pairs also increases until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the new pairs being added at each redshift.

@@

\subsection{Merger Fraction}\label{sec:pairprops-frac}
    We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the total number of merging and non-merging orbits (solid lines) at a given redshift. 
    As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
    Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term.

    Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
    At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
    The merger fraction for both mass ranges declines to zero at $z=0$, due to the very small fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
